---
title: "Lads vs Nups"
output: pdf_document
---

# Comparison of Nup related genes and Robson LADs /  Belmont SPIN classification

```{r,setup,echo=F, message=FALSE, warning=FALSE}

load("lads1_workspace.RData")
library(data.table)
library(ggplot2)
library(dplyr)
library(GenomicRanges)
library(biomaRt)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Robson LADs first

```{r, echo = FALSE,results = F}
## First the Robson data

#we have full coverage here (all Nup associated genes are covered by the Robson DamID data)
# load the Nup related genes from before
nupgenes <- as.data.table(read.csv("C:/Users/38598/Desktop/hivint/nup153/original_associated_genes.csv", sep = "\t"))

# this is messy, lets redo it as a sanity check
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", host    = 'grch37.ensembl.org')
atts <- c(  "ensembl_gene_id",
            "start_position",
            "end_position",
           #"hgnc_symbol",
           "ensembl_transcript_id",
            "ensembl_exon_id",
            "exon_chrom_start",
            "exon_chrom_end",
            "chromosome_name")
ngenes <- getBM(attributes = atts, mart = mart,filters = c("ensembl_gene_id"),
                                                values=unique(nupgenes$ensembl_gene_id))
ngenes <- as.data.table(ngenes)

#  compute introsm
introner <- function(ngenes){
  introns <- ngenes %>%
    arrange(ensembl_transcript_id, exon_chrom_start) %>%
    group_by(ensembl_gene_id, ensembl_transcript_id) %>%
    # pair exons to make introns = gaps between end and next start
    mutate(
      intron_start = lag(exon_chrom_end) + 1,
      intron_end   = exon_chrom_start - 1
    ) %>%
    # drop NA (first exon has no preceding intron)
    filter(!is.na(intron_start) & intron_start <= intron_end) %>%
    dplyr::select(
      ensembl_gene_id,
      ensembl_transcript_id,
      start_position,
      end_position,
      intron_start,
      intron_end,
      chromosome_name
    )

  introns <- as.data.table(introns)
  ngenes[, feature_type := "exon"]
  introns[, feature_type := "intron"]

  # 6) Standardize and select columns to combine
  exon_dt <- ngenes[, .(
    ensembl_gene_id,
    ensembl_transcript_id,
    start_position,
    end_position,
    chromosome_name,
    start  = exon_chrom_start,
    end    = exon_chrom_end,
    feature_type
  )]

  intron_dt <- introns[, .(
    ensembl_gene_id,
    ensembl_transcript_id,
    start_position,
    end_position,
    chromosome_name,
    start  = intron_start,
    end    = intron_end,
    feature_type
  )]

  # 7) Combine into one table
  combined_features <- rbind(exon_dt, intron_dt, fill=TRUE)
  

  # Optional: sort nicely
  setorder(
    combined_features,
    ensembl_gene_id,
    ensembl_transcript_id,
    start
  )
  return(combined_features)
}
ngenes <- introner(ngenes)
rm(combined_features,introns,exon_dt, intron_dt)
ngenes[,chromosome_name := paste0("chr",chromosome_name)]
setkey(ngenes,chromosome_name,start,end)

rladnups <- foverlaps(ngenes,lads)
sum(is.na(rladnups$LAD_binary))  # we miss 42 times, just ignore those
rladnups <- rladnups[!is.na(LAD_binary),]
sum(is.na(rladnups$start))
sum(!ngenes$ensembl_gene_id %in% rladnups$ensembl_gene_id)
# so we have full coverage in the robson data, good

# have to see overlap between integrations in Nup genes and LAD binary to see difference
rladnups[, overlap_len := pmin(end, loc.end) - pmax(start, loc.start) + 1]
rladnups[,lad_width := loc.end - loc.start +1]
rladnups[,feature_width := end - start +1]
rladnups[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of features here!!!
#have to split now for clarity
rladnups_act <- rladnups[type == "Activated",c(1,8,9,10,12,13,14,15,16,17,18,19,20)]
rladnups_rest <- rladnups[type == "Resting",c(1,8,9,10,12,13,14,15,16,17,18,19,20)]

rladnups_act[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]
rladnups_rest[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]
# theres some confusion here with exon and intron ids and different transcript calls for sure

rladnups_act <- unique(rladnups_act)

act_1 = rladnups_act[, .(
  # keep gene width for normalizing
  chromosome_name = unique(chromosome_name),
  start_position = mean(start_position),
  end_position = mean(end_position),
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

rest_1 = rladnups_rest[, .(
  # keep gene width for normalizing
  chromosome_name = unique(chromosome_name),
  start_position = mean(start_position),
  end_position = mean(end_position),
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

act_1$LAD_type <- "Activated Nup"
rest_1$LAD_type <- "Resting Nup"

Robson_LADs_vs_Nups <- rbind(act_1,rest_1)

rm(act_1,rest_1)

# compared to a random selection of genes?
#        wast majority are protein coding genes so compare to that
#   genes table from lads.Rmd was alread protins only
# we can sample it and then get the extra columns fro mbiomart to avoid strain
set.seed(123)
genes_rand <- genes[sample(1:nrow(genes), length(unique(nupgenes$ensembl_gene_id)))]

atts <- c( "ensembl_gene_id",
            "start_position",
            "end_position",
            "ensembl_transcript_id",
            "ensembl_exon_id",
            "exon_chrom_start",
            "exon_chrom_end",
            "chromosome_name")
genes_rand <- getBM(attributes = atts, mart = mart,filters = c("ensembl_gene_id"),
                                                values=unique(genes_rand$ensembl_gene_id))
genes_rand <- as.data.table(genes_rand)
genes_rand <- introner(genes_rand)
genes_rand[,chromosome_name := paste0("chr",chromosome_name)]
setkey(genes_rand,chromosome_name,start_position,end_position)
genes_rand <- foverlaps(genes_rand,lads)

sum(is.na(genes_rand$LAD_binary))  #  great

genes_rand[, overlap_len := pmin(end, loc.end) - pmax(start, loc.start) + 1]
genes_rand[,lad_width := loc.end - loc.start +1]
genes_rand[,feature_width := end - start +1]
genes_rand[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of 
rand_act <- genes_rand[type == "Activated",c(1,8,9,10,12,13,14,15,16,17,18,19,20)]
rand_rest <- genes_rand[type == "Resting",c(1,8,9,10,12,13,14,15,16,17,18,19,20)]

rand_act = rand_act[, .(
  # keep gene width for normalizing
  chromosome_name = unique(chromosome_name),
  start_position = mean(start_position),
  end_position = mean(end_position),
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

rand_rest = rand_rest[, .(
  # keep gene width for normalizing
  chromosome_name = unique(chromosome_name),
  start_position = mean(start_position),
  end_position = mean(end_position),
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]


rand_act$LAD_type <- "Activated RMC"
rand_rest$LAD_type <- "Resting RMC"

Robson_LADs_vs_Nups_RMCs <- rbind(rand_act,rand_rest)

```

### 1. Overlap of Nup related genes and LADs

If there is any overlap between Nup gene and LAD the data point is treated as LAD.
RMCs are randomly chosen protein coding genes with sample size == 486.


```{r, echo = FALSE}
Robson_LADs_vs_Nups[,isLAD := ifelse(pct_LAD > 0, "LAD","Not LAD")]

# the actual plotting code thingy
p1 <- ggplot(Robson_LADs_vs_Nups
, aes(fill=isLAD, x=LAD_type)) + 
  geom_bar(position = position_stack(reverse = TRUE))+ theme_minimal()+
  geom_text(stat = "count", aes(label = ..count..),
            vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "",
       fill = "",
       title = "Any overlap: Nup genes vs LADs") 
p1



# with added RMCs
Robson_LADs_vs_Nups_RMCs[,isLAD := ifelse(pct_LAD > 0, "LAD","Not LAD")]
withRMC <- rbind(Robson_LADs_vs_Nups,Robson_LADs_vs_Nups_RMCs)

p2 <- ggplot(withRMC
, aes(fill=isLAD, x=LAD_type)) + 
  geom_bar(position = position_stack(reverse = TRUE))+ theme_minimal()+
  geom_text(stat = "count", aes(label = ..count..),
            vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "",
       fill = "",
       title = "Any overlap: Nup genes and random genes vs LADs") 
p2

# percent overlap
```

Problem with above graphs: Nup genes are around 2x wider then an average random selection of protein coding genes. This cant explain the full difference but heres the same data but weighted by gene width.
Adjusted overlap score is the number of bp that overlap LAD divided by total length of all genes in group (Nups or RMCs).
The difference is less extreme but still statistically significant.

```{r, echo = FALSE}
plot3 <- withRMC[, .(
  # keep gene width for normalizing
  adjusted_overlap = gene_width * pct_LAD / sum(gene_width)
  # overall percentage of width where there is any LAD overlap
  
), by = c("LAD_type")]

p3 <- ggplot(plot3
, aes(fill = LAD_type,x=LAD_type, y = adjusted_overlap)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "Adjusted overlap score",
       fill = "",
       title = "Adjusted overlap: Nup genes vs LADs")+theme_minimal()+theme(legend.position = "none")
p3

```



### 2. Overview of the interaction between intron/exon split in Nup genes and Robson LADs.

Percentatge overlap thats in exons for genes that have any overlap at all; the Nup gene overlap is mostly intronal: 


```{r, echo = F}
# calculate percentage of the overlap thats exonal
Robson_LADs_vs_Nups[isLAD == "LAD",pct_exon := 100*LAD_exon /(LAD_exon + LAD_intron)]
Robson_LADs_vs_Nups_RMCs[isLAD == "LAD",pct_exon := 100*LAD_exon /(LAD_exon + LAD_intron)]
withRMC2 <- rbind(Robson_LADs_vs_Nups,Robson_LADs_vs_Nups_RMCs)
p4 <- ggplot(withRMC2[isLAD == "LAD",]
, aes(y = (pct_exon), x=LAD_type, fill =LAD_type)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "Percent of exonal overlap",
       fill = "",
       title = "Nup genes overlaping LAD - Intronal pct")+theme_minimal()+theme(legend.position = "none")
p4

```

\newpage

### 3. Comparison of above with HIV integrations

More integrations i Nup genes than random gene sample. Also Robson LADs overlapping genes seem to have small increase in KO153 group integrations.

The recurrent integrations graph isnt adjusted for different group sizes yet (NT is about te same size as KO62 and KO153 together by number of integrations). This is unlikely to explain all of the variance betwen them.

No real difference in percentage of exonal integrations overlapping Activated or Resting LADs, but randomly selected genes have integratiosn fall in more exonal LAD overlaps.


```{r, echo = FALSE}
setkey(Robson_LADs_vs_Nups, chromosome_name, start_position, end_position)
setkey(Robson_LADs_vs_Nups_RMCs, chromosome_name, start_position, end_position)

rlnh_act <- foverlaps(Robson_LADs_vs_Nups[LAD_type == "Activated Nup"],nups[cellType == "Activated"] )
rlnh_rest <- foverlaps(Robson_LADs_vs_Nups[LAD_type == "Resting Nup"],nups[cellType == "Resting"] )

rlnh_act_RMC <- foverlaps(Robson_LADs_vs_Nups_RMCs[LAD_type == "Activated RMC"],nups[cellType == "Activated"] )
rlnh_rest_RMC <- foverlaps(Robson_LADs_vs_Nups_RMCs[LAD_type == "Resting RMC"],nups[cellType == "Resting"] )

# ok now to get something useful out of all of this
plot5 <- rbind(rlnh_act,rlnh_rest,rlnh_act_RMC,rlnh_rest_RMC)
plot5[,hasIntegration:=ifelse(!is.na(sample),"Integration","No integration")]
plot5 <- plot5[!group%in%c("Resting_mock","Activated_mock"), ]

p5 <- ggplot(plot5
, aes(x=isLAD, fill = hasIntegration)) + 
  geom_bar(position  ="fill")+
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+ facet_wrap(~LAD_type)+
  labs(x = "",
       y = "Percent of genes with 1 or more HIV integrations",
       fill = "",
       title = "Nup genes with integration split by Robson LAD status")+theme_minimal()
p5
plot5[is.na(group), group := "No integration"]

p6 <- ggplot(plot5
, aes(x=isLAD, fill = group)) + 
  geom_bar(position  ="fill")+
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+ facet_wrap(~LAD_type)+
  labs(x = "",
       y = "Percent of genes with 1 or more HIV integrations",
       fill = "",
       title = "Nup genes with integration split by Robson LAD status")+theme_minimal()
p6


# do also the total count as well?

plot5[
  , count_in_subgroup := .N, by = .(ensembl_gene_id, group, LAD_type)
]

p7 <- ggplot(plot5[hasIntegration =="Integration",]
, aes(y = count_in_subgroup,x=isLAD, fill = group)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+ facet_wrap(~LAD_type)+
  labs(x = "",
       y = "Number of recurrent integrations in gene",
       fill = "",
       title = "Nup genes with recurrent integration split by Robson LAD status")+theme_minimal()
p7

# consider the Intron exon split
p8 <- ggplot(plot5[,]
, aes(y = log(pct_exon+1),x=LAD_type, fill = group)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+ #facet_wrap(~LAD_type)+
  labs(x = "",
       y = "log(Percent of exonal integrations+1)",
       fill = "",
       title = "Exonal integration percentage split by Robson LAD status")+theme_minimal()
p8

```

\newpage

# Belmont SPIN clasification vs Nup genes

4 nup genes have no overlap with the Belmont Spin classifiction
"CISD1"       "FAM74A2"     "INTS4L2"     "ANKRD20A14P"

Comparison to randomly selected genes shows avoidance of Speckles and first two interior active compartments.

```{r,echo = FALSE, results = FALSE}
setkey(ngenes,chromosome_name,start,end)
tots19$spin_start <- tots19$start
tots19$spin_end <- tots19$end
setkey(tots19,seqnames,spin_start,spin_end)
spinups <- foverlaps(ngenes,tots19) # SPIN vs Nup enes
sum(is.na(spinups$spin)) #750 are not classed
spinups[is.na(spin),spin := "No SPIN overlap"]
#sum(is.na(spinups$start))
#sum(!ngenes$ensembl_gene_id %in% spinups$ensembl_gene_id)
# the missing genes:
#unique(ngenes[which(!ngenes$ensembl_gene_id %in% spinups$ensembl_gene_id),]$ensembl_gene_id)
# so we have almost full coverage in Belmont as well

spinups[, overlap_len := pmin(end, spin_end) - pmax(start, spin_start) + 1]
# define overlap feature starts and stops
spinups[,over_start := pmax(start, spin_start)][,over_end := pmin(end, spin_end)]
spinups[,lad_width := spin_end - spin_start +1]
spinups[,feature_width := end - start +1]
spinups[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of features here!!!

spinups[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]

#RMCS
genes_rand <- genes[sample(1:nrow(genes), length(unique(nupgenes$ensembl_gene_id)))]

atts <- c( "ensembl_gene_id",
            "start_position",
            "end_position",
            "ensembl_transcript_id",
            "ensembl_exon_id",
            "exon_chrom_start",
            "exon_chrom_end",
            "chromosome_name")
genes_rand <- getBM(attributes = atts, mart = mart,filters = c("ensembl_gene_id"),
                                                values=unique(genes_rand$ensembl_gene_id))
genes_rand <- as.data.table(genes_rand)
genes_rand <- introner(genes_rand)
genes_rand[,chromosome_name := paste0("chr",chromosome_name)]
setkey(genes_rand,chromosome_name,start_position,end_position)
genes_rand <- foverlaps(genes_rand,tots19)

sum(is.na(genes_rand$spin)) #750 are not classed
genes_rand[is.na(spin),spin := "No SPIN overlap"]

genes_rand[, overlap_len := pmin(end, spin_end) - pmax(start, spin_start) + 1]
genes_rand[,over_start := pmax(start, spin_start)][,over_end := pmin(end, spin_end)]

genes_rand[,lad_width := spin_end - spin_start +1]
genes_rand[,feature_width := end - start +1]
genes_rand[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of features here!!!

genes_rand[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]

genes_rand$type <- "RMC"
spinups$type <- "Nup gene"

spin_total <- rbind(spinups,genes_rand)
```





```{r,echo = FALSE}
# graphing any of this somehow
p9 <- ggplot(spinups[spin!="No SPIN overlap",]
, aes(fill=spin, x=spin, y= overlap_len)) + 
  geom_col()+ theme_minimal()+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "",
       fill = "",
       title = "Total length of overlap SPIN segments vs Nup genes") +theme(legend.position = "none")
p9



# wight this by total width of each SPIN classification
spinups[,adjusted_overlap := overlap_len / sum(lad_width), by = c("spin","ensembl_gene_id") ]
#spinups[,testing := sum(lad_width), by = c("spin","ensembl_gene_id") ]


p10 <- ggplot(spinups[spin!="No SPIN overlap",]
, aes(fill=spin, x=spin, y= adjusted_overlap)) + 
  geom_col()+ theme_minimal()+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "",
       fill = "",
       title = "SPIN/Nup overlap adjusted by SPIN class size") +theme(legend.position = "none")
p10


p11 <- ggplot(spin_total[spin!="No SPIN overlap",]
, aes(fill=spin, x=spin, y= overlap_len)) + 
  geom_col()+ theme_minimal()+facet_wrap(~type)+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "",
       fill = "",
       title = "Total length of overlap SPIN segments vs Nup genes / RMCs") +theme(legend.position = "none")
p11


spin_total[,adjusted_overlap := overlap_len / sum(lad_width), by = c("spin","ensembl_gene_id","type") ]


p12 <- ggplot(spin_total[spin!="No SPIN overlap",]
, aes(fill=spin, x=spin, y= adjusted_overlap)) + 
  geom_col()+ theme_minimal()+facet_wrap(~type)+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "",
       fill = "",
       title = "SPIN/Nup-RMC overlap adjusted by SPIN class size") +theme(legend.position = "none")
p12

```

\newpage

### Checking if theres any additional integration preference here

KO153 remarkable lack of Interior Act 3 compared to others?

```{r, echo = FALSE}
spin_total2 <- unique(spin_total[,c(1,12,19,27,28,33)])
# this is 958 genes, so RMCs and Nup genes
# remove NAs
spin_total2 <- spin_total2[!is.na(over_start),]
setkey(spin_total2, chromosome_name, over_start, over_end)
# just the genic nups needed
nups2 <- nups[ensembl_gene_id != "Not in gene",]
setkey(nups2,chr_name,integration_start, integration_end)


spinints <- foverlaps(spin_total2,nups2)
spinints <- spinints[!group %in% c("Resting_mock","Activated_mock"),]
spinints[is.na(group),group := "No integration"]

#spinints <- spinints[!is.na(ensembl_gene_id),]
p13 <- ggplot(spinints[spin!="No SPIN overlap",]
, aes(fill=spin, x=spin)) + 
  geom_bar()+ theme_minimal()+facet_grid(rows = vars(type), cols= vars(group))+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  theme(axis.text.x = element_blank())+
  labs(x = "",
       y = "",
       fill = "",
       title = "Integrations vs Nup gene SPIN classes")# +theme(legend.position = "none")
p13




```

# fixing the final graph

```{r, echo = FALSE, results = FALSE}

spinints <- spinints[!ensembl_gene_id == "Not in gene",]
# grab non integrating ones or lose them when dealing with mismatches
noints <- spinints[group == "No integration",]
spinints <- spinints[ensembl_gene_id !=i.ensembl_gene_id ,]
spinints <- rbind(spinints,noints)

p14 <- ggplot(spinints[spin!="No SPIN overlap" & type != "RMC",]
, aes(fill=spin, x=spin)) + 
  geom_bar()+ theme_minimal()+facet_wrap(~group)+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  theme(axis.text.x = element_blank())+
  labs(x = "",
       y = "Number of integrations",
       fill = "",
       title = "Integrations in Nup genes vs SPIN classes")# +theme(legend.position = "none")

p14

p15 <- ggplot(spinints[spin!="No SPIN overlap" & type == "RMC",]
, aes(fill=spin, x=spin)) + 
  geom_bar()+ theme_minimal()+facet_wrap(~group)+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  theme(axis.text.x = element_blank())+
  labs(x = "",
       y = "Number of integrations",
       fill = "",
       title = "Integrations in RMC genes vs SPIN classes")# +theme(legend.position = "none")

p15

p16 <- ggplot(spinints[spin!="No SPIN overlap" & type != "RMC",]
, aes(fill=spin, x=spin)) + 
  geom_bar()+ theme_minimal()+facet_wrap(~group,scales = "free_y")+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  theme(axis.text.x = element_blank())+
  labs(x = "",
       y = "Number of integrations",
       fill = "",
       title = "Integrations in Nup genes vs SPIN classes")# +theme(legend.position = "none")

p16

p17 <- ggplot(spinints[spin!="No SPIN overlap" & type == "RMC",]
, aes(fill=spin, x=spin)) + 
  geom_bar()+ theme_minimal()+facet_wrap(~group,scales = "free_y")+
  #geom_text(stat = "count", aes(label = ..count..),
   #         vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  theme(axis.text.x = element_blank())+
  labs(x = "",
       y = "Number of integrations",
       fill = "",
       title = "Integrations in RMC genes vs SPIN classes")# +theme(legend.position = "none")

p17

# publish the fixed graphs + previpous request

pdf("SPIN_Integrations_v3.pdf", height=6, width=8)
p14
p15
p16
p17
dev.off()

# there is some confusion between the two gene columns due to overlapping genes,
# for example ENSG00000116641 and ENSG00000132855

#saity check to see if 1654 of the 11754 resting NTs really do fall into nup genes
nrow(nups[group == "Resting_NT" & ensembl_gene_id != "Not in gene",])
length(unique(nups[group == "Resting_NT" & ensembl_gene_id != "Not in gene",]$ensembl_gene_id))

sum(unique(nups[group == "Resting_NT" & ensembl_gene_id != "Not in gene",]$ensembl_gene_id) %in% unique(ngenes$ensembl_gene_id))

length(unique(spinints[group == "Resting_NT" & type == "Nup gene",]$i.ensembl_gene_id))
sum(unique(spinints[group == "Resting_NT" & type == "Nup gene",]$i.ensembl_gene_id) %in% nupgenes$ensembl_gene_id)
# solved


# checking again
table(spinints[spin!="No SPIN overlap" & type != "RMC",]$group,spinints[spin!="No SPIN overlap" & type != "RMC",]$spin )

length(unique(spinints[spin!="No SPIN overlap" & type != "RMC" & group == "Resting_NT",]$i.ensembl_gene_id))
# so 268 genes have integrations here
#check if this pans out for Resting_NT specifically
ngenes2 <- unique(ngenes[,c(1,5,3,4)])
setkey(ngenes2,chromosome_name,start_position, end_position)
nups3 <- nups[group =="Resting_NT",]
setkey(nups3, chr_name, integration_start, integration_end)

test <- foverlaps(nups3,ngenes2)
test <- test[!is.na(ensembl_gene_id),]
length(unique(test$ensembl_gene_id))
length(unique(test$uid))

test3 <- test3[!ensembl_gene_id == "Not in gene",]
test3 <- test3[!ensembl_gene_id !=i.ensembl_gene_id]
```

Messing about with LADs in 10% length segments,
with +/1 50% off the ends to see if theres structural length based differences


```{r, echo = FALSE}

# isolate both types of lads lads

rlads <- lads[LAD_binary == "LAD",][,c(2,3,4,8,9)] 
blads <- tots19[,c(1,2,3,4,12,13)]  # lets just keep the whle thing for this one for now
rlads[,width:=loc.end - loc.start +1]
colnames(rlads) <- c("chr","start1","end1","isLAD","cellType","width")
colnames(blads) <- c("chr","start1","end1","width","spin","isLAD")


# gnerate segments


rlads[, w10 := width / 10]
blads[, w10 := width / 10]


offsets <- data.table(
  segment = c(paste0("lead_seg", 1:5),
              paste0("seg", 1:10),
              paste0("trail_seg", 1:5)),
  rel_start = c(-5:-1, 0:9, 10:14),
  rel_end   = c(-4:0, 1:10, 11:15)
)

new_rlads <- rlads[, {
  orig_start <- start1   # preserve original start
  tmp <- copy(offsets)
  tmp[, start := orig_start + rel_start * w10]
  tmp[, end   := orig_start + rel_end   * w10]
  tmp
}, by = .(chr, start1, end1, width, w10, cellType)][
  , c("rel_start","rel_end","w10") := NULL]

new_blads <- blads[, {
  orig_start <- start1   # preserve original start
  tmp <- copy(offsets)
  
  tmp[, start := orig_start + rel_start * w10]
  tmp[, end   := orig_start + rel_end   * w10]
  tmp
}, by = .(chr, start1, end1, width,w10,spin)][
  , c("rel_start","rel_end","w10") := NULL]

# here we messed up since some of Belmonts are on starts of chromosomes
new_blads <- new_blads[start>0,]



library(BSgenome.Hsapiens.UCSC.hg19)
# fix the chromosomal boundry conditions
seqlens <- seqlengths(BSgenome.Hsapiens.UCSC.hg19)

chr_lengths <- data.table(
  chr = names(seqlens),
  chr_len = as.numeric(seqlens)
)

new_rlads <- chr_lengths[new_rlads, on = "chr"]
new_blads <- chr_lengths[new_blads, on = "chr"]
new_rlads[, `:=`(
  start = pmax(1, start),
  end   = pmin(chr_len, end)
)]
new_rlads <- new_rlads[start < end]

new_blads[, `:=`(
  start = pmax(1, start),
  end   = pmin(chr_len, end)
)]
new_blads <- new_blads[start < end]

gr1 <- GRanges(
  seqnames = new_rlads$chr,
  ranges   = IRanges(start = new_rlads$start,
                      end   = new_rlads$end)
)

gr2 <- GRanges(
  seqnames = new_blads$chr,
  ranges   = IRanges(start = new_blads$start,
                      end   = new_blads$end)
)

seqs1 <- getSeq(BSgenome.Hsapiens.UCSC.hg19, gr1)
seqs2 <- getSeq(BSgenome.Hsapiens.UCSC.hg19, gr2)

gc1 <- letterFrequency(seqs1, letters = c("G", "C"), as.prob = TRUE)
gc_content1 <- rowSums(gc1)

gc2 <- letterFrequency(seqs2, letters = c("G", "C"), as.prob = TRUE)
gc_content2 <- rowSums(gc2)

new_rlads[, GC := gc_content1]
new_blads[, GC := gc_content2]

saveRDS(new_rlads,"Segmentated_Robson_LADS.rds")
saveRDS(new_blads,"Segmentated_Belmont_LADS.rds")
```


Since Im separating LADs into percentage based segemnts heres their width distribution:
Theyre pretty uniform so dont worry about it lol

```{r,echo = FALSE}

new_rlads <- readRDS("Segmentated_Robson_LADS.rds")
new_blads <- readRDS("Segmentated_Belmont_LADS.rds")
lads[,width:= loc.end -loc.start +1]
ggplot(lads, aes(x = width)) +
  geom_histogram() 

ggplot(tots19, aes(x = width)) +
  geom_histogram() 

# whatever make the graphs silly billy

head(new_rlads)
setkey(new_rlads,chr,start,end)
setkey(new_blads,chr,start,end,cellType)
#mus tsplit
nra <- new_rlads[cellType == "Activated",]
nrr <- new_rlads[cellType == "Resting",]

setkey(nups,chr_name,integration_start,integration_end)
nra <- foverlaps(nra,nups[cellType =="Activated",])
nrr <- foverlaps(nrr,nups[cellType =="Resting",])
new_rlads <- rbind(nra,nrr)
rm(nra,nrr)
new_rlads <- new_rlads[!is.na(group),]

new_blads <- foverlaps(new_blads,nups)

# filter jsut integrating ones

new_blads <- new_blads[!is.na(group),]

# plot the Rlads

p16 <- ggplot(new_rlads[group %in% c("Activated_KO153","Activated_NT"),], aes(color = group,x = segment, y = GC)) +
  geom_point(alpha = 0.7, size = 1.5, position = "jitter") +# facet_w
  theme_classic() +
  scale_fill_brewer(palette="RdBu", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.8, hjust=0.5))+
  labs(
    x = "Relative Position (10% of segment width increments)",
    y = "GC Content",
    title = "GC content across relative Robson LAD position"
)
p16

# this is probably problematic as some of these overlap other LADS, making a bad image of it all:
  # work on it more!!!!
  # also rather meaningles without comared to RMC integrations


head(new_rlads)

```


using modified ngenes table ( without transcript data) to show how they select into 
various spin categories


```{r}
#saveRDS(ngenes,"nup_genes.rds")
ngenes <- readRDS("nup_genes.rds")
head(ngenes)
ngenes[,chromosome_name:= paste0("chr",chromosome_name)]
setkey(ngenes,chromosome_name,start_position, end_position)
# ok now cross match to 
tots19$spin_start <- tots19$start
tots19$spin_end <- tots19$end
setkey(tots19,seqnames,spin_start,spin_end)

nl <- foverlaps(ngenes,tots19)

nl <- nl[,c(1,12,17,18,19,22)]

n2 <- nl[nl[, .(n_grp1 = uniqueN(spin)), by = ensembl_gene_id][n_grp1 == 3],
   on = "ensembl_gene_id"]

n2 <- unique(n2[,c(1,2,5,6)])
n3 <- unique(nl[,c(1,2,5,6)])


write.csv(nl,"Nup_spin_full.csv")
write.csv(n3,"Nup_spin_shortened.csv")
write.csv(n2,"Nup_spin_unique.csv")


part1 <- nl[, .(grp1_set = unique(spin)), by = ensembl_gene_id][
  , .(
      only_in_subset = all(grp1_set %in% c("Lamina","Near_Lm1","Near_Lm2"))
    ), 
  by = ensembl_gene_id
][only_in_subset == TRUE,ensembl_gene_id]
part11 <- nl[ensembl_gene_id %in% part1,]
part11 <- unique(part11[,c(5,6)])
part11[,spin:="Lamina related"]

part2 <- nl[, .(grp1_set = unique(spin)), by = ensembl_gene_id][
  , .(
      only_in_subset = all(grp1_set %in% c("Interior_Act3"))
    ), 
  by = ensembl_gene_id
][only_in_subset == TRUE,ensembl_gene_id]
part22 <- nl[ensembl_gene_id %in% part2,]
part22 <- unique(part22[,c(5,6)])
part22[,spin:="Interior_Act3"]

nl3 <- rbind(part11,part22)
write.csv(nl3,"Nup_spin_lamina_vs_act3.csv")
```