# Comparison of Nup related genes and Robson LADs /  Belmont SPIN classification

```{r,setup,echo=F, message=FALSE, warning=FALSE}

load("lads1_workspace.RData")
library(data.table)
library(ggplot2)
library(dplyr)
library(GenomicRanges)
library(biomaRt)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## Robson LADs first

```{r, echo = FALSE}
## First the Robson data

#we have full coverage here (all Nup associated genes are covered by the Robson DamID data)
# load the Nup related genes from before
nupgenes <- as.data.table(read.csv("C:/Users/38598/Desktop/hivint/nup153/original_associated_genes.csv", sep = "\t"))

# this is messy, lets redo it as a sanity check
mart <- useMart(biomart = "ensembl", dataset = "hsapiens_gene_ensembl", host    = 'grch37.ensembl.org')
atts <- c(  "ensembl_gene_id",
            "start_position",
            "end_position",
            "ensembl_transcript_id",
            "ensembl_exon_id",
            "exon_chrom_start",
            "exon_chrom_end",
            "chromosome_name")
ngenes <- getBM(attributes = atts, mart = mart,filters = c("ensembl_gene_id"),
                                                values=unique(nupgenes$ensembl_gene_id))
ngenes <- as.data.table(ngenes)

#  compute introsm
introner <- function(ngenes){
  introns <- ngenes %>%
    arrange(ensembl_transcript_id, exon_chrom_start) %>%
    group_by(ensembl_gene_id, ensembl_transcript_id) %>%
    # pair exons to make introns = gaps between end and next start
    mutate(
      intron_start = lag(exon_chrom_end) + 1,
      intron_end   = exon_chrom_start - 1
    ) %>%
    # drop NA (first exon has no preceding intron)
    filter(!is.na(intron_start) & intron_start <= intron_end) %>%
    dplyr::select(
      ensembl_gene_id,
      ensembl_transcript_id,
      start_position,
      end_position,
      intron_start,
      intron_end,
      chromosome_name
    )

  introns <- as.data.table(introns)
  ngenes[, feature_type := "exon"]
  introns[, feature_type := "intron"]

  # 6) Standardize and select columns to combine
  exon_dt <- ngenes[, .(
    ensembl_gene_id,
    ensembl_transcript_id,
    start_position,
    end_position,
    chromosome_name,
    start  = exon_chrom_start,
    end    = exon_chrom_end,
    feature_type
  )]

  intron_dt <- introns[, .(
    ensembl_gene_id,
    ensembl_transcript_id,
    start_position,
    end_position,
    chromosome_name,
    start  = intron_start,
    end    = intron_end,
    feature_type
  )]

  # 7) Combine into one table
  combined_features <- rbind(exon_dt, intron_dt, fill=TRUE)
  

  # Optional: sort nicely
  setorder(
    combined_features,
    ensembl_gene_id,
    ensembl_transcript_id,
    start
  )
  return(combined_features)
}
ngenes <- introner(ngenes)
rm(combined_features,introns,exon_dt, intron_dt)
ngenes[,chromosome_name := paste0("chr",chromosome_name)]
setkey(ngenes,chromosome_name,start,end)

rladnups <- foverlaps(ngenes,lads)
sum(is.na(rladnups$LAD_binary))  # we miss 42 times, just ignore those
rladnups <- rladnups[!is.na(LAD_binary),]
sum(is.na(rladnups$start))
sum(!ngenes$ensembl_gene_id %in% rladnups$ensembl_gene_id)
# so we have full coverage in the robson data, good

# have to see overlap between integrations in Nup genes and LAD binary to see difference
rladnups[, overlap_len := pmin(end, loc.end) - pmax(start, loc.start) + 1]
rladnups[,lad_width := loc.end - loc.start +1]
rladnups[,feature_width := end - start +1]
rladnups[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of features here!!!
#have to split now for clarity
rladnups_act <- rladnups[type == "Activated",c(1,8,9,10,14,15,16,17,18,19,20)]
rladnups_rest <- rladnups[type == "Resting",c(1,8,9,10,14,15,16,17,18,19,20)]

rladnups_act[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]
rladnups_rest[, bp_over_lad := sum(overlap_len), by=c("ensembl_gene_id","feature_type")]
# theres some confusion here with exon and intron ids and different transcript calls for sure

rladnups_act <- unique(rladnups_act)

act_1 = rladnups_act[, .(
  # keep gene width for normalizing
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

rest_1 = rladnups_rest[, .(
  # keep gene width for normalizing
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

act_1$LAD_type <- "Activated Nup"
rest_1$LAD_type <- "Resting Nup"

Robson_LADs_vs_Nups <- rbind(act_1,rest_1)

rm(act_1,rest_1)

# compared to a random selection of genes?
#        wast majority are protein coding genes so compare to that
#   genes table from lads.Rmd was alread protins only
# we can sample it and then get the extra columns fro mbiomart to avoid strain
set.seed(123)
genes_rand <- genes[sample(1:nrow(genes), length(unique(nupgenes$ensembl_gene_id)))]

atts <- c( "ensembl_gene_id",
            "start_position",
            "end_position",
            "ensembl_transcript_id",
            "ensembl_exon_id",
            "exon_chrom_start",
            "exon_chrom_end",
            "chromosome_name")
genes_rand <- getBM(attributes = atts, mart = mart,filters = c("ensembl_gene_id"),
                                                values=unique(genes_rand$ensembl_gene_id))
genes_rand <- as.data.table(genes_rand)
genes_rand <- introner(genes_rand)
genes_rand[,chromosome_name := paste0("chr",chromosome_name)]
setkey(genes_rand,chromosome_name,start_position,end_position)
genes_rand <- foverlaps(genes_rand,lads)

sum(is.na(genes_rand$LAD_binary))  #  great

genes_rand[, overlap_len := pmin(end, loc.end) - pmax(start, loc.start) + 1]
genes_rand[,lad_width := loc.end - loc.start +1]
genes_rand[,feature_width := end - start +1]
genes_rand[,gene_width := end_position- start_position +1] # this is with TSS, so differs from SUM of 
rand_act <- genes_rand[type == "Activated",c(1,8,9,10,14,15,16,17,18,19,20)]
rand_rest <- genes_rand[type == "Resting",c(1,8,9,10,14,15,16,17,18,19,20)]

rand_act = rand_act[, .(
  # keep gene width for normalizing
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]

rand_rest = rand_rest[, .(
  # keep gene width for normalizing
  gene_width = mean(gene_width),
  # overall percentage of width where there is any LAD overlap
  pct_LAD = sum(feature_width * (LAD_binary == "LAD")) / sum(feature_width) * 100,
  # not this ignores TSS for now
  
  LAD_exon  = sum(feature_width * (LAD_binary == "LAD" & feature_type == "exon"))  / sum(feature_width) * 100,
  LAD_intron = sum(feature_width * (LAD_binary == "LAD" & feature_type == "intron")) / sum(feature_width) * 100

), by = ensembl_gene_id]


rand_act$LAD_type <- "Activated RMC"
rand_rest$LAD_type <- "Resting RMC"

Robson_LADs_vs_Nups_RMCs <- rbind(rand_act,rand_rest)

```

1. Overlap of Nup related genes and LADs

If there is any overlap between Nup gene and LAD the data point is treated as LAD.
RMCs are randomly chosen protein coding genes with sample size == 486.
```{r, echo = FALSE}
Robson_LADs_vs_Nups[,isLAD := ifelse(pct_LAD > 0, "LAD","Not LAD")]

# the actual plotting code thingy
p1 <- ggplot(Robson_LADs_vs_Nups
, aes(fill=isLAD, x=LAD_type)) + 
  geom_bar(position = position_stack(reverse = TRUE))+ theme_minimal()+
  geom_text(stat = "count", aes(label = ..count..),
            vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "",
       fill = "",
       title = "Any overlap: Nup genes vs LADs") 
p1



# with added RMCs
Robson_LADs_vs_Nups_RMCs[,isLAD := ifelse(pct_LAD > 0, "LAD","Not LAD")]
withRMC <- rbind(Robson_LADs_vs_Nups,Robson_LADs_vs_Nups_RMCs)

p2 <- ggplot(withRMC
, aes(fill=isLAD, x=LAD_type)) + 
  geom_bar(position = position_stack(reverse = TRUE))+ theme_minimal()+
  geom_text(stat = "count", aes(label = ..count..),
            vjust = -0.3,position = position_stack(reverse = TRUE,vjust = 1)) +
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "",
       fill = "",
       title = "Any overlap: Nup genes and random genes vs LADs") 
p2

# percent overlap
```

Problem with above graphs: Nup genes are around 2x wider then an average random selection of protein coding genes. This cant explain the full difference but heres the same data but weighted by gene width.
Adjusted overlap score is the number of bp that overlap LAD divided by total length of all genes in group (Nups or RMCs).
The difference is less extreme but still statistically significant.

```{r, echo = FALSE}
plot3 <- withRMC[, .(
  # keep gene width for normalizing
  adjusted_overlap = gene_width * pct_LAD / sum(gene_width)
  # overall percentage of width where there is any LAD overlap
  
), by = c("LAD_type")]

p3 <- ggplot(plot3
, aes(fill = LAD_type,x=LAD_type, y = adjusted_overlap)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "LAD cell type",
       y = "Adjusted overlap score",
       fill = "",
       title = "Adjusted overlap: Nup genes vs LADs")+theme_minimal()+theme(legend.position = "none")
p3

```



2. Overview of the interaction between intron/exon split in Nup genes and Robson LADs,
as well as the GC content

Percentatge overlap thats in exons for genes that have any overlap at all; the Nup gene overlap is mostly intronal: 
```{r}
# calculate percentage of the overlap thats exonal
Robson_LADs_vs_Nups[isLAD == "LAD",pct_exon := 100*LAD_exon /(LAD_exon + LAD_intron)]
Robson_LADs_vs_Nups_RMCs[isLAD == "LAD",pct_exon := 100*LAD_exon /(LAD_exon + LAD_intron)]
withRMC2 <- rbind(Robson_LADs_vs_Nups,Robson_LADs_vs_Nups_RMCs)
p4 <- ggplot(withRMC2[isLAD == "LAD",]
, aes(y = (pct_exon), x=LAD_type, fill =LAD_type)) + 
  geom_boxplot()+
  scale_fill_brewer(palette="Blues", direction = 1)+ theme(axis.ticks.y = element_blank())+
  theme(axis.text.x = element_text(angle = 45, vjust = 0.8, hjust=0.5))+
  labs(x = "",
       y = "Percent of exonal overlap",
       fill = "",
       title = "Nup genes overlaping LADs")+theme_minimal()+theme(legend.position = "none")
p4

```

3. Comparison of above with HIV integrations


```{r}

```

\newpage

## Belmont SPIN clasification vs Nup genes

  qqa